---
title: "Covid 19 Infection and Death Rates"
output: html_notebook
---


```{r, include=F}
library(rmarkdown)
library(knitr)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(gridExtra)
library(broom)
```

```{r, include=F}
knitr::opts_chunk$set(
  fig.width=12
)
```



# Load Data from Johns Hopkins
```{r}
covid_infections_raw = read.csv(
  file=url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"),
  header=T,
  stringsAsFactors=F,
  check.names = T
)

covid_deaths_raw = read.csv(
  file=url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"),
  header=T,
  stringsAsFactors=F,
  check.names = T
)
```

# Format data
Data comes with dates as columns and hierarchical index columns. US statistics must be aggregated from individual states, and World statistics by aggregating ALL counts. The transposed data has dates as rows as would be the proper "tidy" format.
```{r}
covid_infections_world = 
  covid_infections_raw %>%
  select(starts_with("X")) %>%
  summarize_all(sum) %>%
  set_rownames('World') %>%
  ### Transpose to make countries the columns
  t() %>%
  data.frame(stringsAsFactors = F) %>%
  rownames_to_column("Date") %>%
  mutate(
    Date =  
      str_remove(Date, "X") %>% 
      as.Date(format="%m.%d.%y")
  ) %>%
  mutate_at(vars(-Date), ~as.numeric(.)) %>%
  mutate(
    days = as.numeric(Date)-as.numeric(min(Date))
  )

covid_deaths_world = 
  covid_deaths_raw %>%
  select(starts_with("X")) %>%
  summarize_all(sum) %>%
  set_rownames('World') %>%
  ### Transpose to make countries the columns
  t() %>%
  data.frame(stringsAsFactors = F) %>%
  rownames_to_column("Date") %>%
  mutate(
    Date =
      str_remove(Date, "X") %>% 
      as.Date(format="%m.%d.%y")
  ) %>%
  mutate_at(vars(-Date), ~as.numeric(.)) %>%
  mutate(days=as.numeric(Date)-as.numeric(min(Date))) 

covid_infections =
  covid_infections_raw %>%
  select(Country.Region, starts_with("X")) %>%
  group_by(Country.Region) %>%
  summarize_all(sum) %>%
  mutate(Country.Region = make.names(Country.Region)) %>%
  column_to_rownames("Country.Region") %>%
  ### Transpose to make countries the columns
  t() %>%
  data.frame(stringsAsFactors = F) %>%
  rownames_to_column("Date") %>%
  mutate(Date =  str_remove(Date, "X") %>% as.Date(format="%m.%d.%y")) %>%
  mutate_at(vars(-Date), ~as.numeric(.)) %>%
  ### Add In World
  inner_join(covid_infections_world, by="Date") %>%
  mutate(
    days = 
      difftime(Date, min(Date), units="days") %>%
      as.numeric()
  )

covid_deaths =
  covid_deaths_raw %>%
  select(Country.Region, starts_with("X")) %>%
  group_by(Country.Region) %>%
  summarize_all(sum) %>%
  mutate(Country.Region = make.names(Country.Region)) %>%
  column_to_rownames("Country.Region") %>%
  ### Transpose to make countries the columns
  t() %>%
  data.frame(stringsAsFactors = F) %>%
  rownames_to_column("Date") %>%
  mutate(Date =  str_remove(Date, "X") %>% as.Date(format="%m.%d.%y")) %>%
  mutate_at(vars(-Date), ~as.numeric(.)) %>%
  ### Add In World
  inner_join(covid_deaths_world, by="Date") %>%
  mutate(
    days = 
      difftime(Date, min(Date), units="days") %>%
      as.numeric()
  )
```

# Plot infection rates
This plot shows a time series of counts since March 1 as well as the entire date range with a log-scale for the y-axis (counts). I chose March 1 because that is the point where the US seemed to enter an exponential phase.
```{r, fig.width=12}
p1 = covid_infections %>%
  filter(Date > as.Date('2020-03-20')) %>%
  select("Date", "US", "Italy", "Spain", "France", "Iran", "World") %>%
  # mutate_at(vars(-Date), ~./max(.)) %>%
  gather(key="Country", value="Infected", -Date) %>%
  ggplot(aes(x=Date, y=Infected, color=Country)) +
  # geom_line() + 
  geom_point(alpha=0.4, size=1.5) +
  scale_y_log10() +
  geom_smooth(method='lm', se=F)  + 
  scale_x_date(date_breaks = "day", date_minor_breaks="day", date_labels="%b %e") 

p2 = covid_infections %>%
  select("Date", "US", "Italy", "Spain", "France", "Iran", "World") %>%
  # mutate_at(vars(-Date), ~./max(.)) %>%
  gather(key="Country", value="Infected", -Date) %>%
  ggplot(aes(x=Date, y=Infected, color=Country)) +
  geom_line() +
  geom_point(alpha=0.4, size=1.5) +
  scale_y_log10()  + 
  scale_x_date(date_breaks = "3 days", date_minor_breaks="day", date_labels="%b %e") 

grid.arrange(p1, p2,
  top = sprintf("Covid 19 Infection Rates -- %s", Sys.Date())
  )
```
# Plot death rates
Same as above but for deaths.
```{r, fig.width=12}
p1 = covid_deaths %>%
  filter(Date > as.Date('2020-03-20')) %>%
  select("Date", "US", "Italy", "Spain", "France", "Iran", "World") %>%
  # mutate_at(vars(-Date), ~./max(.)) %>%
  gather(key="Country", value="Deceased", -Date) %>%
  ggplot(aes(x=Date, y=Deceased, color=Country)) +
  # geom_line() + 
  geom_point(alpha=0.4, size=1.5) +
  scale_y_log10() +
  geom_smooth(method='lm', se=F)

p2 = covid_deaths %>%
  select("Date", "US", "Italy", "Spain", "France", "Iran", "World") %>%
  # mutate_at(vars(-Date), ~./max(.)) %>%
  gather(key="Country", value="Deceased", -Date) %>%
  ggplot(aes(x=Date, y=Deceased, color=Country)) +
  geom_line() +
  geom_point(alpha=0.4, size=1.5) +
  scale_y_log10()

grid.arrange(p1, p2,
  top = sprintf("Covid 19 Death Rates -- %s", Sys.Date())
  )
```

# Calculate Doubling Times
I will use a GLM model to calculate a running 10-day 10-folding and doubling time for both infections and deaths.
## Functions
```{r}
getRunningDoublingTime = function(
  data,
  y,
  from=as.Date(character()),
  to=as.Date(character()),
  lag=0
)
{
  doubling_times = data.frame(
    Date=as.Date(character()),
    y=as.numeric(),
    ymin = as.numeric(),
    ymax = as.numeric()
  )
  
  from = as.Date(from) 
  to = as.Date(to)
  
  # print(sprintf("from: %s  to: %s", from, to))

  for (start in (from-lag+1):(to-lag+2))
  {
    start = as.Date(start, origin="1970-01-01")
    end = as.Date(start+lag-1)
      
    # print(sprintf("Date range: %s to %s", start, end))
    
    formula_string = paste0(y, " ~ as.numeric(Date)")
   
    fit = 
      data %>%
      filter(start <= Date & Date <= end) %>%
      glm(
        formula = as.formula(formula_string),
        family = gaussian(link = "log"), 
        data = .
      )
    
    results = 
      tidy(fit, conf.int=T) %>%
      filter(term=='as.numeric(Date)') %>% 
      mutate_if(is.numeric, ~log(2)/.)
    
    doubling_time = results$estimate
    conf_low = results$conf.high
    conf_high = results$conf.low

    doubling_times %<>%
      add_row(
        Date=end, 
        y=doubling_time,
        ymin=conf_low,
        ymax=conf_high
      )
  }
  
  return(doubling_times)
}
```

## Assemble Data
```{r}
lag = 10

from = "2020-03-10"
# to = "2020-03-27"
to = Sys.Date()

covid_infection_folding = data.frame(
  Date = seq(as.Date(from), as.Date(to), by="day")
)

covid_infection_folding %<>%
  inner_join(
    getRunningDoublingTime(
      covid_infections,
      from=from,
      to=to,
      y="US",
      lag=lag
    ) %>% 
      rename(US_Infections=y, US_Infections_low=ymin, US_Infections_high=ymax) ,
    by="Date"
  ) %>%
  inner_join(
    getRunningDoublingTime(
      covid_infections,
      from=from,
      to=to,
      y="World",
      lag=lag
    ) %>%
    rename(World_Infections=y, World_Infections_low=ymin, World_Infections_high=ymax),
    by="Date"
  ) %>%
  inner_join(
    getRunningDoublingTime(
      covid_deaths,
      from=from,
      to=to,
      y="US",
      lag=lag
    ) %>%
    rename(US_Deaths=y, US_Deaths_low=ymin, US_Deaths_high=ymax),
    by="Date"
  ) %>%
  inner_join(
    getRunningDoublingTime(
      covid_deaths,
      from=from,
      to=to,
      y="World",
      lag=lag
    ) %>%
    rename(World_Deaths=y, World_Deaths_low=ymin, World_Deaths_high=ymax),
    by="Date"
  )
```

# Plot
## US vs World Infections
```{r}
covid_infection_folding %>%
  ggplot() + 
  geom_smooth(aes(x=Date, y=US_Infections, color="US Infection Doubling Time")) +
  geom_smooth(aes(x=Date, y=World_Infections, color="World Infection Doubling Time"))
```

## US vs World Deaths
```{r}
covid_infection_folding %>%
  ggplot() + 
  geom_point(aes(x=Date, y=US_Deaths, color="US Infection Doubling Time")) +
  geom_point(aes(x=Date, y=World_Deaths, color="World Infection Doubling Time")) + 
  geom_smooth(aes(x=Date, y=US_Deaths, color="US Infection Doubling Time")) +
  geom_smooth(aes(x=Date, y=World_Deaths, color="World Infection Doubling Time"))
```

## With Confidence Intervals
### US
```{r}
us_doubling_confint = covid_infection_folding %>%
  ggplot() + 
  geom_point(aes(x=Date, y=US_Infections, color="US Infections")) +
  geom_point(aes(x=Date, y=US_Deaths, color="US Deaths")) +
  geom_smooth(
    data=covid_infection_folding,
    mapping=aes(x=Date, y=US_Infections, ymin=US_Infections_low, ymax=US_Infections_high, color="US Infections"),
    stat="identity"
    ) +
  geom_smooth(
    data=covid_infection_folding,
    mapping=aes(x=Date, y=US_Deaths, ymin=US_Deaths_low, ymax=US_Deaths_high, color="US Deaths"),
    stat="identity"
    ) + 
  scale_x_date(date_breaks = "day", date_minor_breaks="day", date_labels="%b %e") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  labs(
    y="Doubling Time (Days)",
    title="US Covid-19 Doubling Time (95% confidence intervals)",
    subtitle=sprintf("Based on previous %d days", lag)
  )

print(us_doubling_confint)
```

### World
```{r}
world_doubling_confint = covid_infection_folding %>%
  ggplot() + 
  geom_point(aes(x=Date, y=World_Infections, color="World Infections")) +
  geom_point(aes(x=Date, y=World_Deaths, color="World Deaths")) +
  geom_smooth(
    data=covid_infection_folding,
    mapping=aes(x=Date, y=World_Infections, ymin=World_Infections_low, ymax=World_Infections_high, color="World Infections"),
    stat="identity"
    ) +
  geom_smooth(
    data=covid_infection_folding,
    mapping=aes(x=Date, y=World_Deaths, ymin=World_Deaths_low, ymax=World_Deaths_high, color="World Deaths"),
    stat="identity"
    ) + 
  scale_x_date(date_breaks = "day", date_minor_breaks="day", date_labels="%b %e") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  labs(
    y="Doubling Time (Days)",
    title="World COVID-19 Doubling Time (95% confidence intervals)",
    subtitle=sprintf("Based on previous %d days", lag)
  )

print(world_doubling_confint)
```

```{r, fig.width=12}
grid.arrange(
  us_doubling_confint, 
  world_doubling_confint,
  ncol = 2,
  top = sprintf("COVID 19 Doubling Times -- %s", Sys.Date())
  )
```



Using flexdashboard
https://rmarkdown.rstudio.com/flexdashboard/using.html#storyboards

Using flexdashboard
https://rmarkdown.rstudio.com/flexdashboard/using.html#page_icons

Sample flexdashboard Layouts
https://rmarkdown.rstudio.com/flexdashboard/layouts.html#multiple_pages

flexdashboard Examples
https://rmarkdown.rstudio.com/flexdashboard/examples.html

Using flexdashboard
https://rmarkdown.rstudio.com/flexdashboard/using.html#tabsets

1*3Pz0TmNq_gpGVd2xqxfeIQ.png (PNG Image, 1366 × 728 pixels) - Scaled (93%)
https://miro.medium.com/max/3036/1*3Pz0TmNq_gpGVd2xqxfeIQ.png

Building an HR Dashboard in R using Flexdashboard - Towards Data Science
https://towardsdatascience.com/building-an-hr-dashboard-in-r-using-flexdashboard-76d14ed3f32

flexdashboard/htmlwidgets-showcase-storyboard.Rmd at master · rstudio/flexdashboard
https://github.com/rstudio/flexdashboard/blob/master/inst/examples/htmlwidgets-showcase-storyboard.Rmd

Getting started with flexdashboards in R | STATWORX
https://www.statworx.com/at/blog/getting-started-with-flexdashboards-in-r/

Storyboard Showcase
http://rstudio-pubs-static.s3.amazonaws.com/259632_a7c392438653473b8cb0df8f035c4bdb.html

Working on a Workout · Len Kiefer
http://lenkiefer.com/2017/01/18/working-on-a-workout/

Housing’s best year in a decade by @lenkiefer
http://lenkiefer.com/chartbooks/jan2017/review2016sidebar.html#

Housing’s best year in a decade by @lenkiefer
http://lenkiefer.com/chartbooks/jan2017/review2016sidebar.html#

Housing market dashboard by @lenkiefer
http://lenkiefer.com/chartbooks/jan2017/housing-market-dashboard-jan2017.html#
